---
title: 'Lab 11: Checking assumptions for linear regression'
author: "Name and Student ID"
date: "Today's Date"
output: pdf_document
---

* Due date: Friday, November 20th, 11:59 PM.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab11.RAGS.R")
```

```{r load-packages, include = F}
library(broom)
library(dplyr)
library(ggplot2)
library(tidyr)
library(patchwork)
```



### Boston Data on Median Household Value and Distance to Employment Centers

We are examining a data set used to predict housing prices in the area around Boston (Harrison, D. and Rubinfeld, 1978). We wish to specifically examine the association of the measure of housing price (`medv`, median value of owner-occupied homes in the $1000s) and a measure of adjacency to employment (a weighted distance, roughly in miles). The data frame (called 'Boston') is contained in another package (MASS), which we will load below.

```{r, message=F}
### NOTE: All of the code is to get you started on the lab. You do not need to
### understand any functions below that you have not seen before.

# Load the MASS library with the Boston data
library(MASS)

### NOTE: This package has a function `select()` that can be confused with
### dplyr's select. To overcome this, we first import the data we need and then
### detach the library before loading dplyr.

# Load the data
boston_housing <- read.csv("Boston.csv")

# List variables
boston2 <- boston_housing %>% dplyr::select(nox, dis, medv)

# Variable definition - take a quick look at the variables in the data frame

#help(Boston)
detach(package:MASS)

### Normally, when we are doing inference, we take a random sample from the  
### entire population so we can see how well we can make inference when we only 
### have a sample of 50 individuals (rows of data). If you have time after the lab, try taking a random sample of 50 rows from the data, and see if your results change.
```

\newpage

1. Perform a linear regression of `medv` versus `nox` using the Boston data and summarize the results. Be careful about which variable is explanatory and 
which is response!

```{r}
# write your code here. Make a linear model, ~tidy~ the output, and report the slope

p1 <- "Write the slope coefficient here rounded to 2 decimals"


p1
check_problem1()
```

2. Interpret the slope parameter for `nox`. Notice that the other columns are `std.error`, `statistic`, and `p-value` -- these should remind you of things we've learned about in inference in Part III of the course. They correspond to the hypothesis test with the null hypothesis that the parameter is equal to 0. Thus, how would you interpret the p-value for `nox`?

[Write your answer here.]



3. Use `glance()` to look at the $r^2$ value for this model. Does `nox` explain 
a lot of the variation in median household value? Would you expect it to?

```{r}
#Put your code here
```

[Write your answer here.]



4. Check the assumptions required for the simple linear model using the plots
shown during lecture. Note that to make these plots you need to first fit the linear model and then use the `augment()` function from the broom package to store the residuals and fitted values into a new data frame.

The hardest plot to make is likely the boxplot because the data first needs to 
be reshaped. You can reshape the data with the `gather()` function that you see in the slides. Here is a helpful explanation for how gather() works: https://twitter.com/WeAreRLadies/status/1059520693857996800.

Basically, we need to gather the observed y values and the residuals by stacking
them into one variable so that we can make two box plots side-by-side. Below,
we include the `gather()` code for you since it is a bit tricky. You need to use
the resulting data frame to make your box plots.

```{r }
# Put your code here

# augment your model

# first plot
plot1 <- "Your plot"

# second plot
plot2 <- "Your plot"

# third plot
plot3 <- "Your plot"

# fourth plot (gather code included for you. It assumes your augmented data is
# called `augmented_1`, so you will likely need to update that to whatever your
# augmented data is called.)

#reshape <- augmented_1 %>% dplyr::select(.resid, medv) %>%
#gather(key = "type", value = "value", medv, .resid)

plot4 <- "Your plot"

plot1 
plot2
plot3
plot4

#NO AG for this question
```

5. What do you think about the assumption plots? They appear to be a bit messier than the ones shown in class, but these reflect what we often see "in the real world".

[Write your answer here.]



--------------------------------------------------------------------------


### Lab Conclusion (make sure to read this and understand it)**

From this exercise, we can conclude that there is a negative association between
nitrogen oxides and median household value. An increase of 1 part per 10 million (PPM) of nitrogen oxide is associated with a decrease in median household value 
of \$33,900 (see `help(Boston)` to remind yourself of the units for `nox` and 
`medv`). Note that this "increase of 1 unit" is wider than what we see in the range of the scatterplot, so we should modify our interpretation to reflect a 0.1 unit increase in `nox`. In other words, an increase in Nitrogen Oxide of 0.1 PPM is associated with a decrease in median household value of \$3,390. This is easier to visualize when you look at the scatterplot of the data and the line of best fit. Look at the increase from 0.5 to 0.6 on the x-axis and see how the model predicts a decrease in the household value from ~\$25k to ~\$22k.

\newpage


6. Perform a linear regression of `medv` (median value of owner-occupied homes in \$1000s) and `dis` (weighted mean of distances to five Boston employment centers) using the Boston data and summarize the results. 

Assign the linear model to an object called `p6`.

Be careful about which variable is explanatory and which is response! 

```{r p1}
# write your code here.
p6 <- "YOUR CODE HERE"

p6
check_problem6()
```

\newpage

7. Interpret the slope parameter and p-value from the table. What null and alternative hypotheses does this p-value refer to?  

[Write your answer here.]



\newpage

8. Derive a 95% CI for this slope parameter and assign the object `p8` to a vector of the lower bound and upper bound of the interval. Round to AT LEAST one decimal place. In your opinion, would you expect the direction of this relationship to hold if the data were collected today?

```{r p8}
#Put your code here
p8 <- "YOUR ANSWER HERE"

p8
check_problem8()
```

[Write your answer here.]


\newpage

9. Use a function to look at the r-squared value for this model. Round the r-squared value to 2 decimal places and assign this value to the object `p9`. Does `dis` explain a lot of the variance in median household value? Would you expect it to? 

```{r p4}
#use the function here

p9 <- "put r-squared value here"

p9
check_problem9()
```

[Write your answer here.]


\newpage

10. Make a plot with the raw data points, the fitted line from the simple linear regression model (only containing `medv` and `dis`), and add a line with a slope of 0. You can have the horizontal line cross the y-axis at the average value of `medv` to vertically bisect the data points. Store your plot as the object `p10`.

```{r p5, echo=F, fig.align='center', out.width="80%"}
p10 <- "CODE YOUR PLOT HERE"


check_problem10()
```

\newpage

11. Does your plot raise any concerns about the assumptions of the linear regression?  What other plots might you create to explore the fit of the model? One helpful plot would compare the distribution of model residuals to a theoretical normal distribution. Assign the object `p11` to the FIRST TWO LETTERS of the name of this plot.

```{r p6}
p11 <- "YOUR ANSWER HERE"
### OPTIONAL: CODE THE PLOT

p11
check_problem11()
```

[Write your answer here.]


\vspace{20pt}

Regardless of your answer, we go forward using the model to make inferences about the points on the line.

\newpage

### Pointwise Confidence Intervals and Multiple Testing

As you learned in lecture, there are two types of confidence intervals applicable to estimating a point on the plot which are related to whether one is predicting the population average among individuals with $X=x$ (**mean response**) or whether one is predicting the actual $Y$ for a particular individual (**single observation**). For this assignment, we will concentrate on the confidence interval for the mean response. We do so because it is rare to use statistical models in public health as forecasting models (predicting an individual's health in the future) and more common to use them to estimate population-level changes (how does the mean health change in a population as we change exposure). However, as precision medicine becomes more of a reality and the models accurately predict health (i.e., have high $R^2$'s), then statistical forecasting may become more common in our field.

12. Calculate four 95% confidence intervals for the mean response, one at each `dis` value: 2.5, 5.0, 7.5, and 10.0 miles. Store the lower bounds for each confidence interval, ROUNDED to two decimal places, in a vector called `p12`.

**Hint: Use the predict function, and be sure to specify interval = "confidence"**

OPTIONAL: If time allows, add the four CIs to a scatter plot of the data (along with the line of best fit).

```{r p7}
#Put your code here
### Helpful Data Frame:
ci_dataframe <-data.frame(dis = c(2.5, 5.0, 7.5, 10))
p12 <- "YOUR CODE AND ANSWER HERE"

p12
check_problem12()
```

\newpage

13. Interpret the pointwise 95% confidence interval of the median house price when distance = 10.  

[Write your answer here.]



\newpage

14. Do the CI's differ in length for different values of `dis`?  Why or why not?

[Write your answer here.]




\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.

```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-fa20/lab/lab11; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
