---
title: 'Lab 10: Chi Squared Testing'
author: "Name and Student ID"
date: "Today's Date"
output: pdf_document
---

* Due date: Friday, Nov 13th 23:59pm.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab10.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

### Instructions


1) We will be using data tidying functions and plotting functions to work through this lab.  Load the 
required packages into our R session.

```{r Setup}
```

### Chi Squared Testing

As the textbook mentions, the chi-square statistic is a measure of how far 
the observed counts in a two-way table are from the expected counts. The 
formula for the statistic is:

$$ X^2 = \sum\frac{(count_{observed} - count_{expected})^2}{count_{expected}} $$

The sum is over all cells in the table. That is, there are as many terms in 
the sum as there are cells in the table. Each term in the sum is called a 
$X^2$ component.

**Part 1: Melanoma**
Adapted from Baldi and Moore Question 21.29

Melanoma is a rare form of skin cancer that accounts for the great majority of skin cancer fatalities. UV exposure is a major risk for melanoma. A question we would like to explore is if the body parts which have increased sun exposure are more susceptible to melanoma. A random sample of 310 women diagnosed with melanoma were classified according to the known location of the melanoma on their bodies. Here are the results:

Location | Head/Neck | Trunk | Upper Limbs | Lower Limbs |
---------|-----------|-------|-------------|-------------|
Count    |     45    |   80  |     34      |      151    |
---------|-----------|-------|-------------|-------------|
Expected |    XXXX   |  XXXX |     XXXX    |     XXXX    |
---------|-----------|-------|-------------|-------------|

1. Assuming each of the four locations represent roughly equal skin areas, fill in the expected counts for the four areas of the body.  

[Your answer here]



2. What are the assumptions for completing a Chi Squared test? Are the conditions met for this example?

[Your answer here]



3. Perform a chi-squared test for goodness of fit. State the null and alternative hypotheses. Use R to calculate your test statistics and report this value. Calculate and report the p-value. 

[Your answer here]  

$H_0$ : Your null hypothesis
$H_A$ : Your alternative hypothesis


```{r Q1 Melanoma Locations}
#Replace with your code/work.

test_stat <- "Your test statistic"
p_value <- "Your calculated p-value here (round to nearest whole number)"


test_stat
p_value

check_problem3()
```


\newpage

**Part 2**
Adapted from: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_HypothesisTesting-ChiSquare/BS704_HypothesisTesting-ChiSquare2.html

The National Center for Health Statistics (NCHS) provided data on the distribution of weight (in categories) among Americans in 2002. The distribution was based on specific values of body mass index (BMI).

Underweight was defined as BMI< 18.5, Normal weight as BMI between 18.5 and 24.9, overweight as BMI between 25 and 29.9 and obese as BMI of 30 or greater. 

Americans in 2002 were distributed as follows: 2% Underweight, 39% Normal Weight, 36% Overweight, and 23% Obese. Suppose we want to assess whether the distribution of BMI is different in the Framingham Offspring sample.

Using data from the n = 3,326 participants who attended the seventh examination of the Offspring in the Framingham Heart Study we created the BMI categories as defined and observed the following:

BMI      | Underweight | Normal Weight | Overweight  |     Obese   | Total |
---------|-------------|---------------|-------------|-------------|-------|
Count    |     20      |     932       |    1374     |    1000     |  3326 |
---------|-------------|---------------|-------------|-------------|-------|
Expected |    XXXX     |    XXXXXX     |    XXXXXX   |     XXX     | 3326 |
---------|-------------|---------------|-------------|-------------|-------|

4. State the null and alternative hypotheses. Fill in the expected counts for this statistic.  

[Your answer here]  

$H_0$:
$H_A$:



5. For this test, we will use a 5% signficance level. For what value of the test statistic under Chi-Squared distribution will we reject the null hypothesis?
(Hint: What are the degrees of freedom for this test?)

```{r Question 2 BMI part b}
#Replace with your code
reject_value <- "Your answer here round to 4 decimals"


reject_value

check_problem5()
```

6. Perform a chi-squared test for goodness of fit. Calculate and report your test statistic. Calculate and interpret your p-value. What are your conclusions?

```{r Question 2 BMI part c}
#Replace with your code


```


[Your answer here]




\newpage

### Chapter 22 - Chi Squared Test for Independence


The chi-square test for a two-way table with r rows and c columns uses
critical values from the chi-square distribution with (r-1)(c-1) degrees
of freedom.

* Side question: Think about how might we determine a p-value for a chi-square test statistic?

In research, we are often interested in making the assumption that two explanatory variables are (mostly)
independent.  Independence is actually one condition which permits us to include multiple explanatory variables 
in a linear regression (i.e. the line of best fit model that you explored in the first part of the course). 
Thus, the Chi-Square test of independence can be quite useful as a tool to explore whether two 
categorical variables show substantial dependence.

In the second part of this lab, we proceed to walk through the data cleaning, visualization, and 
analysis required to carry out a Chi Square test for two-way tables.


\newpage


## Part 3. Intro and Data are from the text (Ex 22.40 Do angry people have more heart disease??):

*NOTE: If at any point, you are unclear how to use dplyr to create a variable, feel free to 
manually calculate, and use the following code to add a manual variable:

```{r}
# chd_by_anger_level <-
#   chd_by_anger_level %>%
#   ### input the 6 values below
#   mutate(new_variable = c( , , , , , ))
```


People who get angry easily tend to have more heart disease. That’s the conclusion of a study that 
followed a random sample of 12,986 people from three locations for about four years. 
All subjects were free of heart disease at the beginning of the study. 
The subjects took the Spielberger Trait Anger Scale test, which measures how prone a person 
is to sudden anger. Here are data for the 8474 people in the sample who had normal blood pressure. 
18 CHD stands for “coronary heart disease.” This includes people who had heart attacks and 
those who needed medical treatment for heart disease.

|        | Low anger   | Moderate anger | High anger |
|--------|-------------|----------------|------------|
| CHD    |     53      |      110       |     27     |
| No CHD |     3057    |     4621       |    606     |


Let's explore if these data support the conclusion of the study!


7. Based on the two-way table above and the framework of the study, write out the null and alternative 
hypotheses that we will be exploring via a Chi-Squared test.


[Your answer here]



We need to figure out what the expected counts of heart disease and anger levels would be if the 
two categorical variables are independent.  Here is a data set of our two-way table values:

```{r Create Dataset, include=FALSE}
chd_by_anger_level <- 
  data.frame(anger_level   = c("Low", "Low",    "Moderate", "Moderate", "High", "High"),
             heart_disease = c("CHD", "No CHD", "CHD",      "No CHD",   "CHD",  "No CHD"),
             actual_count  = c(53,    3057,     110,        4621,       27,     606),
             stringsAsFactors = FALSE)
```


8. Using our dplyr tools, add variables for row, column totals, and sample size, and fill out the two-way table below:
[HINT: The code for computing row totals is given.  Use this framework to compute column totals]  
[new variable names: row_total, column_total, sample_size]
```{r Totals}
### Code your answer here. You will have to repeat this syntax a few times.
chd_by_anger_level <-
  chd_by_anger_level %>%
  group_by(heart_disease) %>%
  mutate(row_total = sum(actual_count)) %>%
  ungroup()


check_problem8()
  
```



|              | Low anger   | Moderate anger | High anger |  Row Total    |  
|--------------|-------------|----------------|------------|---------------|
| CHD          |    XX       |      XXX       |     XX     |     XXX       |
| No CHD       |     XXXX    |     XXXX       |    XXXX    |    XXXX       |
|--------------|-------------|----------------|------------|---------------|
| Column Total |    XXXX     |      XXXX      |   XXX      |    XXXX       |


9. Use the following formula from lecture notes to create a column for expected counts:
$$E_i =\frac{\text{row total}\times\text{col total}}{\text{overall total}}$$

```{r Expected counts}
# add expected counts (a new variable named "expected_count") to your dataframe here


check_problem9()
```



10. Before moving forward with analysis, confirm two requirements for the Chi-Squared test of independence, namely:

* No more than 20% of the expected counts are smaller than 5.0, and 
* All individual expected counts are 1.0 or greater.

```{r 3d Scratch Work}
# [Your answer here]

```
[Your answer here]


While, we are set to move forward with a Chi-Square test, let's practice visualizing our data to see if there 
may be a significant relationship between heart disease and anger.

11. First, calculate the probability of anger level conditional on CHD Status

```{r Calculated conditional distribution}
# add conditional probabilities (a new variable named "cond_prop_anger") to your dataframe here


check_problem11()
```


12. Reference your Ch 22 notes and create a dodged 
bar graph of anger probabilities, dodged by CHD status.  Interpret the results.

```{r Create Plot}
dodged <- "Your plot code here"

dodged

check_problem12()
```


Write your interpretation here.

Now, we are ready to move forward with our Chi-Square test of independence.

13. Compute the Chi-Square test statistic. 
[Optional: Practice dynamic coding. Assign important variables to your environment once, and only call the variable names when computing the 
final test statistic.]


```{r Computing the Test Statistic}
# if you choose to use dynamic coding, you can do it here

chi_square_test_statistic <- "Your statistic"

chi_square_test_statistic

check_problem13()
```
If you choose to calculate your statistic by hand, show your work here:



14. Determine the p-value of your Chi-Square test statistic and interpret the p-value for 
a 5% level Chi-Square test in the context of this problem.

```{r Determine p-value}
chi_sq_p_value <- "Your p-value"


chi_sq_p_value
check_problem14()
```

15. How might we have tested for independence between anger and heart disease prevalence during the 
probability section of the course?  Would we have found that anger and heart disease are independent 
using our old problem-solving process?  How does this method differ from comparing conditional 
probabilities?

[YOUR ANSWER HERE]


\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-fa20/lab/lab10; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.
