---
title: 'Lab 7: Introducing the Central Limit Theorem and Confidence Intervals'
output: pdf_document
---

* Due date: Friday, October 16 10:00pm.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab07.RAGS.R")
```

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage


# (Optional) Relevant Textbook Exercises
The following questions in your textbook are helpful practice for 
understanding today's material:

Baldi and Moore: Ex. 13.5, 13.8, 13.9, 13. 10, 13.12, 13.14


# Introduction 

You will use a central data source (a Google sheet) attached to the lab to
develop a very concrete idea of sampling distributions, and to see the central 
limit theorem in action.  

# The underlying population

Suppose you had a data frame containing the **entire population** of all
residents of Alameda County. You have data on three variables:

1.  Born either out (=1) versus in (=0) the county.
2.  Number of siblings (integer)
3.  Number of visits to the hospital last year

1. Read in the data, L06_Alameda.csv, and assign it to the name `alameda_pop`. Calculate the true 
(population) mean, and make histograms or bar charts of the distribution for each 
variable.  

```{r readin, message = F, warning = F}
library(dplyr)
library(ggplot2)
library(readr)

#### YOUR CODE GOES HERE ####

```

Notice that the distribution of number of siblings and number of hospital visits
are discrete distributions and both skewed right. Today, we will focus on the 
distribution of the number of siblings `num_sibs`. Remember, we know the population
mean exactly, because we have all the data. You calculated the underlying population
mean in the code chunk above.

### Calculating the sampling distribution

We are now going to look at the approximate **sampling distribution for the sample mean**
of the `num_sibs` in (live) action. Remember from earlier lectures that a 
**sampling distribution** is a distribution for a statistic, like the sample 
proportion or the sample mean.

EEach student will be tasked with repeatedly taking a random sample of the population. 
Once you take your sample you will compute the sample mean and add it to a 
shared google sheet. As the data is added to the google sheet, notice how the graph 
to the right of the data changes; this illustrate how the sampling distribution varies for 
increasingly larger sample sizes.

The GSIs will provide you the link to the communal google sheet. The columns in 
the sheets are  `n (Sample_size)` `mean(numSibs)` `Name (your sign in)`.

### Your task

1. Randomly generate 10 simple random samples of size $n=5$ from the population.
Calculate the average number of siblings for each of your samples. We wrote code
to start you off, which you can simply copy and paste 10 times to generate ten 
randomly drawn samples and their sample means.

One sample has been provided to you. Try it yourself 10 times and record your results in the vector below.


```{r size5}
#### YOUR CODE GOES HERE ####

#repeat this code 10 times.
size_5 <- sample_n(alameda_pop, 5, replace = FALSE)
size_5 %>% summarise(mean_numSibs=mean(num_sibs))

```

After you've calculated 10 sample means using the above code, copy and paste your
data into the google sheet for your lab section. The links to all the 
google sheets are:

- 101B (Thursday 5-7pm): https://docs.google.com/spreadsheets/d/130FOigRrcdzyMpxSmzxEcdx2TOthUeKZFBpfJgHpbWk/edit?usp=sharing
- 102B (Wednesday 2-4pm): https://docs.google.com/spreadsheets/d/1sVx2Vd_57DWNHRvQ1UP_BhaMuRUnc-H-L-f5Ujuv-fs/edit?usp=sharing
- 103B (Friday 9-11am): https://docs.google.com/spreadsheets/d/1sJpPRIyl83CcFwOyfwJCgyM1LubDZx-RvAfYwG1v8iA/edit?usp=sharing
- 104B (Wednesday 9-11am): https://docs.google.com/spreadsheets/d/1iOUF5bohUL3_tEHkjOjXC_bueFVLm-sA-lWtxHFLgS0/edit?usp=sharing
- 105B (Wednesday 4-6pm):  https://docs.google.com/spreadsheets/d/1PwcGfK4dKbybxFfrwXm8-H4yCzKVBzCenDaf902Btlg/edit?usp=sharing
- 106B (Wednesday 5-7pm): https://docs.google.com/spreadsheets/d/1AP3oewsuzDDpG6oPnMEqXXjzYNY1Q2EBQjycb5aSokg/edit?usp=sharing
- 107B (Thursday 8-10am):  https://docs.google.com/spreadsheets/d/15JdSP4V3-K5BY76sxRXL3wQcDWhNkuAXxPXzqD88Fo0/edit?usp=sharing
- 108B (Wednesday 1-3pm):  https://docs.google.com/spreadsheets/d/151RrhBvC33sb_4mJDl5ibNOPJRKvKQK4LlCCHg-1pDA/edit?usp=sharing
- 109B (Wednesday 9-11am): https://docs.google.com/spreadsheets/d/1bR2eXPNrdKIi_Cxw5nllQFPwR3-KWcvhA93a6IfRxhc/edit?usp=sharing
- 110B (Friday 9-11am):  https://docs.google.com/spreadsheets/d/17Td3U8t4MoH0nCOTevhJjsiENPaYAWT_egRclsxM9IM/edit?usp=sharing

Once the sheet is full, look at the plot of the **sampling distribution** for
the mean number of siblings when $n=5$. 


- What is the range of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]



Once the class has examined the sampling distribution when $n=5$, repeat the 
same steps for n=50. 


2. Repeat the above process for a sample size of $n=50$.

```{r size50}
#### YOUR CODE GOES HERE ####


size_50_samples <- c("YOUR VALUES HERE")

```

After you calculated your 10 sample means, navigate to the google sheet from 
before, but switch to the sheet with $n=50$ (you can switch tabs in the bottom left).
Add your data for $n=50$.

Look at the plot to the right on the google sheet; now with $n=50$

- What is the range of the sampling distribution of the mean? How does it compare
to when $n=5$?

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? How does it compare to when $n=5$? 

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean? 

[TODO: YOUR ANSWER HERE]


3. Repeat the entire process for sample size $n=500$.

```{r size500}
#### YOUR CODE GOES HERE ####

size_500_samples <- c("YOUR VALUES HERE")

```

- What is the range of the sampling distribution of the mean? How does it compare to when $n=5$ and $n=50$?

[TODO: YOUR ANSWER HERE]


- What is the shape of the sampling distribution of the mean? How does it compare
to when $n=5$ and $n=50$?

[TODO: YOUR ANSWER HERE]


- What is the mean of the sampling distribution of the mean?

[TODO: YOUR ANSWER HERE]



Suppose you have 500 classmates, and they have done this lab and added their data to this 
sheet: https://docs.google.com/spreadsheets/d/1AXStOd618raoWvrBequxOh5CDwgisFJoHo50fmcKb_E/edit?usp=sharing

Open the link, and look at the resulting sampling distributions for $n=5$, $n=50$, and $n=500$.
This is what happens when you repeat the sampling 5,000 times.


4. For which sample size should the sampling distribution of the mean be most normal?
  
  (a) n=5
  (b) n=50
  (c) n=500

Assign your letter choice as a string. Example: sampleSize_answer<-"b"
```{r}
sampleSize_answer<- "REPLACE WITH a,b, or c. Keep the quotes"
sampleSize_answer


check_problem4()

```

\newpage

## Part 2: Confidence intervals

Calculate the true (population) mean of the variable `height`, and the population standard deviation. To do this, use a dplyr function to make a dataframe object called `height_mean_sd` with the first column called `mean_height` and the second column called `sd_height`.

Make a histogram called `plot1` of `height` and comment its distribution. Use ggplot() to quickly
make a ggplot. Does it look Normally distributed?

```{r, message = F, warning = F}
height_mean_sd <- "Your code here"
plot1 <- "Your code here"

height_mean_sd
plot1

check_problem5()
```

[TODO: Your answers here]

Suppose that you know the population standard deviation. It is what you calculated
in the previous code chunk. Use the following code chunk to save its value. Replace
the word `NULL` with your calculation of the population standard deviation.:

```{r save-known-pop-sd}
known_pop_sd <- NULL


check_problem6()
```

Now, suppose you *do not know* the population mean, but wanted to estimate it 
based on a sample. In this lab, we actually know the true value because we 
calculated it above. This way, we can see how well any one sample does at estimating
the population mean and see how often the confidence intervals contain the mean
across several repeated samples.

### Calculating the CI and looking at its performance

We are now going to compute (and enter into our master google sheet) 95%
confidence intervals (CI) for sampling means of different sizes. For this lab we:

- Have a variable with an underlying Normal distribution
- Will take simple random samples (SRS) from this distribution
- Know the value of the population mean (from your calculation in the first code
chunk)

Thus, we satisfy the three conditions discussed in Wednesday's lecture for calculating
a confidence interval when the underlying SD is known.

Recall the format for the 95% confidence interval in this setting (Hover with 
your mouse within the double-dollar signs to see the formula or knit the file to 
read them easily):

$$\bar{x} \pm 1.96 \times\frac{\sigma}{\sqrt{n}}$$
Where:

- The sample mean is the estimated mean, based on your sample
- $\sigma$ is the known standard deviation `known_pop_sd`, that you saved earlier for the distribution of `height`
- $\sigma/\sqrt{n}$ is the standard error of the sampling distribution for $\bar{x}$
- 1.96 is the critical value for a 95% CI.

Like Part 1, each student will take a few samples from the distribution. This
time, you will calculate the mean of your sampled heights and its confidence
interval using the above formula. We will then record this information into the 
google sheet and plot all the CIs when we have at least 20 of them.

Here are the links to the communal Google sheets. Open the one for your lab.
The columns in the sheets are `sample_mean_heights` `lower_CI` `upper_CI` 
`sample_size` `student`.

Lab 101B: https://docs.google.com/spreadsheets/d/1bWatXMlpJLvOR0ch4Hox-5kMHcsit-32nAuPXhYgXMI/edit?usp=sharing
Lab 102B: https://docs.google.com/spreadsheets/d/1V8TtSV6uv7prZIYcGQTn9QQ482IajVYLX6E4LZjwfRw/edit?usp=sharing
Lab 103B: https://docs.google.com/spreadsheets/d/1diY17Q78CVczyLFOSwlySRwF0Ef9T96o8IYO6lHEjPA/edit?usp=sharing
Lab 104B: https://docs.google.com/spreadsheets/d/1gVGxDOPrlk3jvKtI5raezOj17WF1IhJaxt3gs1KR4Ds/edit?usp=sharing
Lab 105B: https://docs.google.com/spreadsheets/d/1sYpg1MkTcIR2ViHdQZrtVJ7_oR5OUwVJT-f_LwmBjQo/edit?usp=sharing
Lab 106B: https://docs.google.com/spreadsheets/d/1TMpkxrOXvfp7-u6KxnQiIRFvUuIdrFDNDl6hHwSqtS0/edit?usp=sharing
Lab 107B: https://docs.google.com/spreadsheets/d/15V6qR37-pPW0t3DAAuH_ElyqToTBgtFp0DYLu3iR0Og/edit?usp=sharing
Lab 108B: https://docs.google.com/spreadsheets/d/1TFQYLia1HG1GA6hPGy1WwoFYS-qav5fvcbshwrqD7zs/edit?usp=sharing
Lab 109B: https://docs.google.com/spreadsheets/d/1Axn4186utxbIUGcKu_SNp_u2pd7rtTPNsSlv4HTmvps/edit?usp=sharing
Lab 110B: https://docs.google.com/spreadsheets/d/1pwv2WJbpy_oZZT7D5TN0Gb8-vh-sFx5SkK60Fkr51AI/edit?usp=sharing

### Your task

1. Randomly generate 3 simple random samples of size $n = 10$ from the population.
Calculate the average number of siblings for each of your samples. We wrote code
to start you off, **but you need to replace the three instances of NULL with 
calculations** to compute the sample mean (`sample_mean_heights`), the lower 
confidence interval (`lower_CI`) and the upper confidence interval (`upper_CI`).
Hint: Review the above section for tips on how to calculate the CI if you forget.
Once you do this, you can simply copy and paste 3 times to generate three 
randomly-drawn samples and their sample means.

```{r sample-size-10}
sample_size <- 10
critical_value <- 1.96
size_10 <- sample_n(alameda_pop, sample_size, replace = FALSE)
size_10 %>% summarise(mean_heights = NULL) %>%
mutate(lower_CI = NULL,
       upper_CI = NULL
)
```

Navigate to the google sheet for your lab and add the mean and its CI for your
three samples. Once this is done enough times, the GSI can make a plot of the
CIs and see how many contain the true value for the mean. Based on this plot:

- What proportion of the confidence intervals contain the mean? 

Type your answer here.

2. Repeat for a sample size of $n=50$. In the code chunk below, generalize your
code from the previous chunk to create three samples, this time of size 50:

```{r sample-size-50}
# add in your code here.
```

After you calculated your 3 sample means, navigate to the Google sheet from 
before, but switch to the sheet with $n=50$ (you can switch tabs in the bottom left).
Add your data for $n=50$.

Once this is done, the GSI will plot these data, now with $n=50$

- What proportion of the confidence intervals contain the mean?

Type your answer here.

- How do the average widths of the CI's compare for $n=50$ versus $n=10$?

Type your answer here.

- What would happen to the confidence intervals if $n=500$?

Type your answer here.


\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-fa20/lab/lab07; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

