---
title: "Lab 3: Relationship between global cesarean delivery rates and GDP"
author: "Chandler B. 303..."
date: "September 9, 2020"
output: pdf_document
---

```{r setup, include = FALSE}
# Don't change these lines, just run them!
source("setup/lab03.RAGS.R")

```


### Instructions 
* Due date: Friday, September 11 at 11:59pm PST.
* Late penalty: 50% late penalty if submitted within 24 hours of due date, no 
marks for assignments submitted thereafter.
* This assignment is graded on **correct completion**, all or nothing. You must pass all public tests and submit the assignment for credit.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.


Start by loading the required libraries, reading in the data and adding on a 
variable:

```{r initial-setup}
library(dplyr)
library(ggplot2)
library(readr)
library(broom)

CS_data <- read_csv("cesarean.csv")

# This code re-orders the variable Income_Group in the specified order.
# Note that it *does not* change the order of the data frame (like arrange() does)
# Rather, it specifies the order the data will be plotted. 
# This will make more sense when we plot the data using Income_Group, and then
# again using Income_Group_order
CS_data$Income_Group <- forcats::fct_relevel(CS_data$Income_Group, 
                                            "Low income", "Lower middle income", 
                                            "Upper middle income", "High income: nonOECD", 
                                            "High income: OECD")

CS_data <- CS_data %>% mutate(CS_rate_100 = CS_rate*100)
```

\newpage 

**1. [1 point] Make a scatter plot between `CS_rate_100` and `GDP_2006`:**

```{r scatterplot}
p1 <- ggplot(CS_data, aes(x = GDP_2006, y = CS_rate_100)) +
  geom_point()
p1


check_problem1()
```

In your plot, you might notice that many of the points are condensed towards the 
lower left corner. And you might recall from the lab and assignment that the 
distributions of both cesarean delivery rate and GDP covered a wide range of 
values. Both of these variables are good candidates for log transformations to
spread out the range of data at the lowest levels.

\newpage

**2.[1 point]  Using the `mutate()` function, add two new logged variables to the data set
`CS_data` and assign this new data set to `CS_data_log`. Call the variables `log_CS` and `log_GDP`. Use base e, also known as natural logarithms, to create the logged variables:**

```{r create-logged variables}
CS_data_log <- CS_data %>% mutate(log_CS = log(CS_rate_100), log_GDP = log(GDP_2006))


check_problem2()
```
\newpage

**3. [1 point] Remake the scatter plot using the logged variables**:

```{r scatter-logged}
p3 <- ggplot(CS_data_log, aes(x = log_GDP, y = log_CS)) +
  geom_point()
p3


check_problem3()
```
\newpage

**4. [1 point]  A `geom` that you have not yet learnt is `geom_smooth()`. This geom can fit 
a curve to the data. Extend your `ggplot()` code by adding geom_smooth() to it:**
```{r add-smoothed-curve}
p4 <- ggplot(CS_data_log, aes(x = log_GDP, y = log_CS)) +
  geom_point() +
  geom_smooth()
p4

check_problem4()
```

**Does the relationship between logged GDP and logged CS look linear?**

The relationship between logged GDP and logged CS is linear for the first 60% of our data. The points start to plateau after a log_GDP of 8.


\newpage 

**5. [1 point] Modify your scatter plot by linking the color of the points to the variable
`Income_Group`.**

```{r color-points-by-income-group}
p5 <- ggplot(CS_data_log, aes(x = log_GDP, y = log_CS)) +
  geom_point(aes(col = Income_Group)) +
  geom_abline(slope = 0.457, intercept = -1.412) +
  geom_smooth()
p5

check_problem5()
```

**Based on this colored scatter plot, it looks like the relationship is linear 
for those countries that are not categorized as one of the two high income
categories.**

```{r}
lm(log_CS ~ log_GDP, CS_data_log)
```


\newpage

**6. [1 point] For this lab, we would like to use linear regression. To do this, use a `dplyr` function to make a new data set called `CS_data_sub` that only contains the low-, lower-middle, and upper-middle income countries (hint: You might want to look at the data to see exactly what these levels are called in the data set):**

```{r subset-low-mid-income}
CS_data_sub <- CS_data_log %>% filter(Income_Group %in% c("Low income", "Lower middle income", "Upper middle income"))

#CS_data_sub <- CS_data_log %>% filter(!(Income_Group %in% c("High income: nonOECD", "High income: OECD")))


check_problem6()
```

\newpage

**7. [1 point] Remake the last scatter plot, this time using `CS_data_sub` to see if the 
relationship looks approximately linear between the logged variables:**

```{r scatter-low-mid}
p7 <- ggplot(CS_data_sub, aes(x = log_GDP, y = log_CS)) +
  geom_point(aes(col = Income_Group)) +
  geom_smooth()
p7


check_problem7()
```

**8. [1 point] Given that the relationship is approximately linear, use linear regression to 
model the relationship between `log_CS` as the response variable and `log_GDP` 
as the explanatory variable. Don't forget to specify the correct data set!:**

```{r linear-model}
p8 <- lm(formula = log_CS ~ log_GDP, data = CS_data_sub)
tidy(p8)

check_problem8()
```

**9.Interpret the slope estimate:**

A one unit increase in the natural log of GDP is associated with a 0.82 unit increase in the natural log of the cesarean delivery rate.


**10.Estimate what the cesarean delivery rate would be for a country with a GDP of 2000. Outline the steps you take to calculate your answer and provide an interpretation. Round your final answer to one decimal place.**

We first need to convert the GDP onto the log (base $e$) scale. We can use the 'log()' function in R to take the logarithm of 2000 using base $e$. To use base 10, we can use the log10() function instead.

logCS = logGDP * slope + Intercept

* Step 1: Take the natrual log of GDP of 2000: log(2000) = 7.600902
* Step 2: Plug the logged GDP into the line of best fit formula:
    - logCS = logGDP * slope + Intercept
    - logCS = 7.600902 * slope + Intercept
    - logCS = 7.600902 * (0.819335) - 3.9404856
    - logCS = 2.287199
* Step 3: Exponentiate both sides of the equation to get back to the cesarean delivery rate.
    - exp(logCS) = exp(2.287199)
    - CS_rate = 9.847317% = 9.8%

Interpretation: A country with a GDP of 2000 is expected to have a cesarean delivery rate of 9.8%



**11. Is it appropriate to use the model to predict the cesarean delivery rate for a country with a GDP of 50,000? Why or why not? Based on the relationship in the full data set, would you expect the linear model to over or under predict?**

A GDP of 50,000 corresponds to a logged GDP of 10.82. This value is outside of the range of our data that was used for the linear model. Because of this, we should NOT use the model to make this prediction because then we would be extrapolating data. If we did use the model, it would likely over-predict the country's cesarean delivery rate because the relationship between GDP and cesarean delivery rate is less steep for high income countries (more accurately, it plateaus after a log_GDP of around 8).


\newpage

### Check your score

Click on the middle icon on the top right of this code chunk (with the downwards gray arrow and green bar) to run all your code in order. Then, run this chunk to check your score.
```{r check-total-score}
# Just run this chunk.
total_score()
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-fa20/lab/lab03; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. 
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza under the post "Submission Issues". 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

